---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!--README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/cfjdoedens/drawsneeded/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cfjdoedens/drawsneeded/actions/workflows/R-CMD-check.yaml)

[![Codecov test
coverage](https://codecov.io/gh/cfjdoedens/drawsneeded/graph/badge.svg)](https://app.codecov.io/gh/cfjdoedens/drawsneeded)

<!-- badges: end -->

```{r init, message=TRUE, warning=TRUE, include=FALSE, paged.print=TRUE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.path = "man/figures/README-",
out.width = "100%"
)
```

# drawsneeded

## Introduction

Supports planning of statistical sampling to audit the defect rate in a set of like items.
This package is useful, when the purpose of the audit is to establish whether the 
defect rate of the total set with e.g. .95 certainty is below a threshold, 
the maximally allowed defect rate.
drawsneeded() assumes the defect rate of an individual item is expressed 
on a scale of 0 tot 1. 

An example application is the audit of a file of monetary statements.
In general this concerns money that has
been spent. Here the defect rate is the fraction of money from the file
that was wrongly spent. And the defect rate of an individual element 
typically varies from 0 to 1, with probably most elements having 0 defect
rate and some elements having substantial defect rates.

Another example application is the audit of a file of administrative decisions
that have to be checked on their legality.
Here the defect rate is the fraction of decisions in the file 
that are not legally correct.
Typically the defect of an individual item is either 0 or 1.

This package has the function 

* drawsneeded(posited_defect_rate, allowed_defect_rate, cert)

The function
gives an estimate of the number of drawn items needed to
establish with some certainty, _cert_, that the defect rate is below a certain
threshold, _allowed_defect_rate_. The assumption of drawsneeded() is
that each drawn item exactly hase _posited_defect_rate_. 
It might be that in practice the items have a defect
of either 0 or 1, or, when they have a defect 
on a scale from 0 to 1, most items have 0 defect and most items will contain no 
defects at all, and some items
will have a large defect rate, or are even totally defective.
Still, we suppose that this assumption leaves the estimates
of drawsneeded() realistic.

Due to randomness, the actual number of
needed drawn items might be smaller or bigger. 
Also, it might turn out, that the set has a higher defect rate than 
allowed_defect_rate. In that case no amount of draws will suffice
to prove that the defect rate is below allowed_defect_rate.

The package has a couple of plotting functions:

* drawsneeded_plot()
* margin_plot_varying_posited_defect_rate()
* margin_plot_varying_allowed_defect_rate()
* margin_plot_varying_cert()
* combined_plot()

These should give the user insight in the relation between number
of draws taken and resulting estimates about the defect rate in the
set of items that is audited. 

## When can drawsneeded() be useful?

drawsneeded() can be useful in the following situation:

1. you want to audit the defect rate of a set of like items, and
1. you consider doing this audit by means of statistical sampling 
   on the set of items, and
1. you want to establish whether or not the defect rate is
   below a certain treshold, and
1. you want to estimate how many items you need to sample

## Role of posited_defect_rate

As said, the function drawsneeded(posited_defect_rate, allowed_defect_rate, cert)
gives an estimate of the minimum number of drawn items needed to
establish with some certainty, cert, e.g. 0.95, that the defect rate is below allowed_defect_rate.
For this estimation it assumes that the actual defect rate in the file
of items equals the posited_defect_rate. 
It is up to the user of drawsneeded() to choose a
posited_defect_rate().
When the posited_defect_rate is chosen conservative, i.e.
on the high side, it can be used with some confidence to produce a not too
low number. And vice versa we can set the posited_defect_rate on the low side
to get a bottom estimate of the number of monetary statements that need 
to be drawn and checked.

The function assumes that each drawn item has a defect rate of posited_defect_rate.
And it computes the minimum number of drawn items necessary to establish with cert certainty
that posited_defect_rate < allowed_defect_rate. Note that it might actually not
be possible for any of the items in the file to have exactly 
that defect rate.

The computation of this minimum is based on the binomial distribution. This is extended
with the beta distribution to allow for non integer values of defects.
(The binomial distribution assumes that there are very many items to draw from, which
all have equal chance of being drawn, and there are so many items that when we have drawn some
this does not change in a substantial way the rate of erroneous items.)

Note that in this scheme the statistical procedure that is used to estimate
the defect rate in the file of items does _not_ use
the posited_defect_rate! The posited_defect_rate is only used as means
to estimate the minimal number of items that need to be drawn to estimate
the defect rate with sufficient certainty.
The posited_defect_rate is used only for planning.

## Usage

### General

Using drawsneeded() is a simple and cheap way to estimate
the effort, needed to statistically establish the 
defect rate in a set of like items. 

### No need to stick to the planning for actual drawing and checking

The number that is produced by drawsneeded() gives an estimate of
how many drawn items are needed to establish the defect rate in a set of 
like items. In order to have a statistical sound
estimate of the defect rate, it is not necessary to stick to that planned number
for actual drawing and evaluation!
The number is just an indication of the effort needed.

For example, when drawsneeded() estimates that there are 365
draws needed (see the first example below), it might
turn out that after 298 draws checked, not one defect was
found. If the 298 items were chosen randomly, 
over the total
set of items, it can be safely concluded that the defect rate is not
more than 1 %. And there is no need to check the other 67 items.

On the other hand, when it turns out that the defect rate of the
drawn and evaluated so far items is above the posited_defect_rate,
one can do a new call on drawsneeded() to establish a new, and 
forcibly higher, number of draws that are needed. Care should be taken that all
items in the file had, and have, an equal chance of being selected.

### Combination with incremental drawing

In order to minimize the number of to be drawn and to be checked
items one can explicitly proceed by making an optimistic,
i.e. low, guess of the posited_defect_rate, and once this indeed turns
out to be too optimistic, plan anew as described above with
the actual defect rate found in the already checked items.

## Installation

You can install the development version of drawsneeded from
[GitHub](https://github.com/) with:

```{r install, message=TRUE, warning=TRUE, echo=TRUE}
if (file.exists("/home/crist-jan/R/x86_64-pc-linux-gnu-library/4.5/drawsneeded")) {
  # We are executing on the author machine, use the development version available there.
  library("drawsneeded")
  print("using files directly from author of package; not from github")
} else {
  # Use the github version.
  if (!requireNamespace("pak", quietly = TRUE)) {
    install.packages("pak")
  }
  pak::pak("cfjdoedens/drawsneeded")
  print("using files from github")
}
```
## Example: 0.1 percent defect expected

### posited_defect_rate = *0.001*, allowed_defect_rate = 0.01, cert = 0.95

Suppose you know from previous experience that a small defect rate might
exist in the set of items. You guess the defect rate to
be no more than 0.1 percent for the total set. So you set
posited_defect_rate to 0.001. You need a result with 95% certainty. So
you set certainty to 0.95.

```{r tiny-defects-num}
drawsneeded(posited_defect_rate = 0.001, allowed_defect_rate = 0.01, cert = 0.95)
```

The conclusion is that you will need at least 365 samples.

In a picture this looks like:

```{r tiny-defects-plot}
drawsneeded_plot(posited_defect_rate = 0.001,  allowed_defect_rate = 0.01, cert = 0.95)
```
We can further analyse the situation by varying over posited_defect_rate, or over allowed_defect_rate or over cert.
See the following three plots.

### Varying over the posited defect rate

First we vary over the posited defect rate:

```{r tiny-defects-vary-eer-plot}
margin_plot_varying_posited_defect_rate(allowed_defect_rate = 0.01, cert = 0.95, max_n = 1000)
```

We see that as posited_defect_rate goes near to allowed_defect_rate, the
number of draws needed rises. In this graph only the results up to 1000 draws are shown.
This limit can be made higher, but this does not really give a more interesting picture:


```{r tiny-defects-vary-eer-plot-100000}
margin_plot_varying_posited_defect_rate(allowed_defect_rate = 0.01, cert = 0.95, max_n = 100000)
```

We can also study the variation of posited_defect_rate in a non graphical way:

```{r tiny-defects-vary-eer-num}
drawsneeded(posited_defect_rate = seq(from = 0.0, by = 0.001, to = 0.009), allowed_defect_rate = 0.01, cert = 0.95)
```

Again we see that for a posited defect rate of 0.001, we need 365 draws.

### Varying over the allowed defect rate

```{r tiny-defects-vary-aer-plot}
margin_plot_varying_allowed_defect_rate(posited_defect_rate = 0.001, cert = 0.95)
```

We see that as the
allowed_defect_rate moves away from the posited_defect_rate, the number of
draws falls sharply.

Non graphically this looks like (note the 365 computed draws needed for allowed defect rate 0.01):

```{r tiny-defects-vary-aer-num1}
drawsneeded(posited_defect_rate = 0.001, allowed_defect_rate = seq(from = 0.002, by = 0.001, to = 0.01), cert = 0.95)
```

and,

```{r tiny-defects-vary-aer-num2}
drawsneeded(posited_defect_rate = 0.001, allowed_defect_rate = seq(from = 0.02, by = 0.01, to = 0.1), cert = 0.95)
```
and,
```{r tiny-defects-vary-aer-num3}
drawsneeded(posited_defect_rate = 0.001, allowed_defect_rate = seq(from = 0.2, by = 0.1, to = 0.9), cert = 0.95)
```

### Varying over the certainty

Finally we can also vary over the certainty we apply.

```{r tiny-defects-vary-cert-plot}
margin_plot_varying_cert(posited_defect_rate = 0.001, allowed_defect_rate = 0.01)
```

As might be expected: if we can do with less certainty, then we need less draws.

In raw numbers this looks like:

```{r tiny-defects-vary-cert-num}
drawsneeded(posited_defect_rate = 0.001, allowed_defect_rate = 0.01, cert = seq(from = 0.50, by = 0.05, to = 0.95))
```

Note that for cert 0.95, we get again 365 draws needed.

### All plots bundled in one picture

We can call combined_plots() to get all four plots in one picture. However, it does not look
very slick here due to lack of canvas space.

```{r tiny-defects-all-plots}
combined_plots(posited_defect_rate = 0.001, allowed_defect_rate = 0.01, cert = 0.95)
```



## Example: no defect expected

### posited_defect_rate = *0*, allowed_defect_rate = 0.01, cert = 0.95

You might expect to see no defect at all in the to be audited set of items. Then you could
calculate as follows:

```{r zero-defects-num}
drawsneeded(posited_defect_rate = 0, allowed_defect_rate = 0.01, cert = 0.95)
```

So then 298 samples is sufficient.

In a picture this looks like:

```{r zero-defects-plot}
drawsneeded_plot(posited_defect_rate = 0, allowed_defect_rate = 0.01, cert = 0.95)
```

Again, we can further analyse the situation by varying over posited_defect_rate, or over allowed_defect_rate or over cert. As we have above already studied varying the posited_defect_rate,
we will not repeat that here.

### Varying over the allowed defect rate

```{r zero-defects-vary-aer-plot}
margin_plot_varying_allowed_defect_rate(posited_defect_rate = 0, cert = 0.95)
```

We see that as the
allowed_defect_rate moves away from the posited_defect_rate, the number of
draws falls sharply.

Non graphically this looks like:

```{r zero-defects-vary-aer-num1}
drawsneeded(posited_defect_rate = 0, allowed_defect_rate = seq(from = 0.001, by = 0.001, to = 0.01), cert = 0.95)
```

and,

```{r zero-defects-vary-aer-num2}
drawsneeded(posited_defect_rate = 0, allowed_defect_rate = seq(from = 0.02, by = 0.01, to = 0.1), cert = 0.95)
```
and,
```{r zero-defects-vary-aer-num3}
drawsneeded(posited_defect_rate = 0, allowed_defect_rate = seq(from = 0.2, by = 0.1, to = 0.9), cert = 0.95)
```

### Varying over the certainty

Finally we can also vary over the certainty we apply.

```{r zero-defects-vary-cert-plot}
margin_plot_varying_cert(posited_defect_rate = 0, allowed_defect_rate = 0.01)
```

As might be expected: more certainty needs more draws.

In raw numbers this looks like:

```{r zero-defects-vary-cert-num}
drawsneeded(posited_defect_rate = 0, allowed_defect_rate = 0.01, cert = seq(from = 0.50, by = 0.05, to = 0.95))
```

We recognize the number 298 as the number of draws needed for cert 0.95.

### All plots bundled in one picture

We call combined_plots() to get all four plots in one picture. Again, it is does not look
very nice here, because of the cramped space. 

```{r zero-defects-all-plots}
combined_plots(posited_defect_rate = 0, allowed_defect_rate = 0.01, cert = 0.95)
```



## Suggested extensions

-   Handle case where only integer values of k are possible. For example
    when k \> 1 - cert, then round k up. Otherwise round k down.
-   In extension to the above:
    Handle the case where per audited item only defect rates of fixed size are possible.
    E.g. an item has as only possible defect rates: 0.25, 0.5, 0.75 and 1.0.
-   Add extra margin, so extra draws, by taking into account
    the standard deviation.
-   Add a prior to drawsneeded().
-   Add a cost function that makes additional draws more expensive.
    The function can help with planning of incremental drawing.
