---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!--README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/cfjdoedens/drawsneeded/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cfjdoedens/drawsneeded/actions/workflows/R-CMD-check.yaml)

[![Codecov test
coverage](https://codecov.io/gh/cfjdoedens/drawsneeded/graph/badge.svg)](https://app.codecov.io/gh/cfjdoedens/drawsneeded)

<!-- badges: end -->

```{r init, message=TRUE, warning=TRUE, include=FALSE, paged.print=TRUE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.path = "man/figures/README-",
out.width = "100%"
)
```

# drawsneeded

## Introduction

Supports monetary unit sampling. In monetary unit sampling we have a
file of monetary statements. In general this concerns money that has
been spent. We want to estimate the percentage of money from the file
that was wrongfully spent, the error rate.

This package has the function 

* drawsneeded(expected_error_rate, allowed_error_rate, cert, max_n)

The function
gives an estimate of the number of monetary unit draws needed to
establish with some certainty, _cert_, that the error rate is below a certain
threshold, _allowed_error_rate_. The assumption of drawsneeded() is
that each drawn statement contains 100 x _expected_error_rate_ percent
error. Most often, in real life, most monetary statements will contain no errors at all, and some statements
will have large errors, or are even totally wrong.
Still, we suppose that this assumption leaves the estimates
of drawsneeded() realistic.

Due to randomness, the actual number of
monetary unit draws needed might be smaller or bigger. 
Also, it might turn out, that the file has a higher error rate than 
allowed_error_rate. In that case no amount of draws will suffice
to prove that the error rate is below allowed_error_rate.

The package has a couple of plotting functions:

* drawsneeded_plot()
* margin_plot_varying_expected_error_rate()
* margin_plot_varying_allowed_error_rate()
* margin_plot_varying_cert()
* combined_plot()

These should give the user insight in the relation between number
of draws taken and resulting estimates about the error rate in the
file that is audited. 

## Interpretation of the value returned by drawsneeded()

As said, the function drawsneeded(expected_error_rate, allowed_error_rate, cert, max_n)
gives an estimate of the number of monetary unit draw items needed to
establish with some certainty, cert, e.g. 0.95, that the error rate is below allowed_error_rate.
For this estimation it assumes that the actual error rate in the file
of monetary statements equals the expected_error_rate.

Of course, this assumption is nearly always wrong as it is highly
unlikely that the expected error rate equals exactly the actual error rate.
However, when the expected_error_rate is chosen conservative, i.e.
on the high side, it can be used with some confidence to produce a not too
low number.

The function then assumes that each drawn item has an error rate of expected_error_rate.
And it computes the minimum number of drawn items necessary to establish with cert certainty
that expected_error_rate < allowed_error_rate.

The computation of this minimum is based on the binomial distribution. This is extended
with the beta distribution to allow for non integer values of errors.
(The binomial distribution assumes that there are very many items to draw from, which
all have equal chance of being drawn, and there are so many items that when we have drawn some
this does not change in a substantial way the rate of erroneous items.)

The idea is that we can in this way estimate the number of to be drawn items, 
to statistically prove that the actual error rate is smaller than the allowed error rate.
And this without using in the proof any knowledge about the expected error rate.
To summarize, the function uses the expected error rate to compute an estimate 
of the necessary number of to be drawn items, where only these checked
items are used to prove the value of the error rate of the file of monetary items.
So expected_error_rate is only used for the planning part!

## Usage

### General

Using drawsneeded() is a simple and cheap way to estimate
the effort, needed to statistically establish the 
error rate in a file of financial statements. 

### No need to stick to the planning for actual drawing and checking

The number that is produced by drawsneeded() gives an estimate of
how many draws are needed to establish the error rate in a file of 
financial statements. In order to have a statistical sound
estimate of the error rate of a file of
financial statements it is not necessary to stick to that planned number
for actual drawing and evaluation!
The number is just an indication of the effort needed.

For example, when drawsneeded() estimates that there are 365
draws needed (see one of the examples below), it might
turn out that after 298 draws checked, not one error was
found. If the 298 items were chosen randomly, based on monetary value,
over the total
file of monetary statements it can be safely concluded that the error rate is not
more than 1 %. And there is no need to check the other 67 items.

On the other hand, when it turns out that the error rate of the
drawn and evaluated so far items is above the expected_error_rate,
one can do a new call on drawsneeded() to establish a new, and 
forcibly higher, number of draws that are needed. The extra
items to be drawn should be chosen randomly chosen, based on monetary value,
over the total file of monetary statements.

### Combination with incremental drawing

In order to minimize the number of to be drawn and to be checked
monetary statements one can explicitly proceed by making an optimistic,
i.e. low, guess of the expected_error_rate, and once this indeed turns
out to be too optimistic, plan anew as described above with
the actual error rate found in the already checked monetary statements.

## Installation

You can install the development version of drawsneeded from
[GitHub](https://github.com/) with:

```{r install, message=TRUE, warning=TRUE, echo=TRUE}
if (file.exists("/home/crist-jan/R/x86_64-pc-linux-gnu-library/4.5/drawsneeded")) {
  # We are executing on the author machine, use the development version available there.
  library("drawsneeded")
  print("using files directly from author of package; not from github")
} else {
  # Use the github version.
  if (!requireNamespace("pak", quietly = TRUE)) {
    install.packages("pak")
  }
  pak::pak("cfjdoedens/drawsneeded")
  print("using files from github")
}
```
## Example: 0.1 percent error expected

### expected_error_rate = *0.001*, allowed_error_rate = 0.01, cert = 0.95, max_n = 500

Suppose you know from previous experience that a small error rate might
exist in the mass of monetary statements. You estimate the error rate to
be no more than 0.1 percent of the total mass in money. So you set
expected_error_rate to 0.001. You need a result with 95% certainty. So
you set certainty to 0.95, You decide that the maximum number of samples
you are willing to draw and check is 500. So you set max_n to 500.

```{r tiny-errors-num}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

The conclusion is that you will need at least 365 samples.

In a picture this looks like:

```{r tiny-errors-plot}
drawsneeded_plot(expected_error_rate = 0.001,  allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

We can further analyse the situation by varying over expected_error_rate, or over allowed_error_rate or over cert.
See the following three plots.

### Varying over the expected error rate

First we vary over the expected error rate:

```{r tiny-errors-vary-eer-plot}
margin_plot_varying_expected_error_rate(allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

We see that as expected_error_rate goes near to allowed_error_rate, the
number of draws needed rises.

We can also study this in a non graphical way:

```{r tiny-errors-vary-eer-num}
drawsneeded(expected_error_rate = seq(from = 0.0, by = 0.001, to = 0.009), allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

Again we see that for an expected error rate of 0.001, we need 365 draws.

### Varying over the allowed error rate

```{r tiny-errors-vary-aer-plot}
margin_plot_varying_allowed_error_rate(expected_error_rate = 0.001, cert = 0.95, max_n = 500)
```

We see that as the
allowed_error_rate moves away from the expected_error_rate, the number of
draws falls sharply.

Non graphically this looks like (note the 365 computed draws needed for allowed error rate 0.01):

```{r tiny-errors-vary-aer-num1}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = seq(from = 0.002, by = 0.001, to = 0.01), cert = 0.95, max_n = 500)
```

and,

```{r tiny-errors-vary-aer-num2}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = seq(from = 0.02, by = 0.01, to = 0.1), cert = 0.95, max_n = 500)
```
and,
```{r tiny-errors-vary-aer-num3}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = seq(from = 0.2, by = 0.1, to = 0.9), cert = 0.95, max_n = 500)
```

### Varying over the certainty

Finally we can also vary over the certainty we apply.

```{r tiny-errors-vary-cert-plot}
margin_plot_varying_cert(expected_error_rate = 0.001, allowed_error_rate = 0.01, max_n = 500)
```

As might be expected: if we can do with less certainty, then we need less draws.

In raw numbers this looks like:

```{r tiny-errors-vary-cert-num}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = seq(from = 0.50, by = 0.05, to = 0.95), max_n = 500)
```

Note that for cert 0.95, we get again 365 draws needed.

### All plots bundled in one picture

We can call combined_plots() to get all four plots in one picture. However, it is does not look
very slick here due to lack of canvas space.

```{r tiny-errors-all-plots}
combined_plots(expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```



## Example: no error expected

### expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500

You might expect to see no error at all in the to be audited file of monetary statements. Then you could
calculate as follows:

```{r zero-errors-num}
drawsneeded(expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

So then 298 samples is sufficient.

In a picture this looks like:

```{r zero-errors-plot}
drawsneeded_plot(expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

Again, we can further analyse the situation by varying over expected_error_rate, or over allowed_error_rate or over cert. As we have above already studied varying the expected_error_rate,
we will not repeat that here.

### Varying over the allowed error rate

```{r zero-errors-vary-aer-plot}
margin_plot_varying_allowed_error_rate(expected_error_rate = 0, cert = 0.95, max_n = 500)
```

We see that as the
allowed_error_rate moves away from the expected_error_rate, the number of
draws falls sharply.

Non graphically this looks like:

```{r zero-errors-vary-aer-num1}
drawsneeded(expected_error_rate = 0, allowed_error_rate = seq(from = 0.001, by = 0.001, to = 0.01), cert = 0.95, max_n = 500)
```

and,

```{r zero-errors-vary-aer-num2}
drawsneeded(expected_error_rate = 0, allowed_error_rate = seq(from = 0.02, by = 0.01, to = 0.1), cert = 0.95, max_n = 500)
```
and,
```{r zero-errors-vary-aer-num3}
drawsneeded(expected_error_rate = 0, allowed_error_rate = seq(from = 0.2, by = 0.1, to = 0.9), cert = 0.95, max_n = 500)
```

### Varying over the certainty

Finally we can also vary over the certainty we apply.

```{r zero-errors-vary-cert-plot}
margin_plot_varying_cert(expected_error_rate = 0, allowed_error_rate = 0.01, max_n = 500)
```

As might be expected: more certainty needs more draws.

In raw numbers this looks like:

```{r zero-errors-vary-cert-num}
drawsneeded(expected_error_rate = 0, allowed_error_rate = 0.01, cert = seq(from = 0.50, by = 0.05, to = 0.95), max_n = 500)
```

We recognize the number 298 as the number of draws needed for cert 0.95.

### All plots bundled in one picture

We call combined_plots() to get all four plots in one picture. Again, it is does not look
very nice here, because of the cramped space. 

```{r zero-errors-all-plots}
combined_plots(expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```



## Suggested extensions

-   Handle case where only integer values of k are possible. For example
    when k \> 1 - cert, then round k up. Otherwise round k down.
-   In extension to the above:
    Handle the case where per audited monetary statement only error rates of fixed size are possible.
    E.g. a monetary statement has as only possible error rates: 0.25, 0.5, 0.75 and 1.0.
-   Add extra margin, so extra draws, by taking into account
    the standard deviation.
-   Add a prior to drawsneeded().
-   Add a cost function that makes additional draws more expensive.
    The function can help with planning of incremental drawing.
