---output:github_document---< !--README.md is generated from README.Rmd. Please edit that file- ->
```{r init, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.path = "man/figures/README-",
out.width = "100%"
)
```

# drawsneeded

< !--badges:start- ->
< !--badges:end- ->

The pacakage drawsneeded supports monetary unit sampling.
In monetary unit sampling we have a file of monetary statements.
In general this concerns money that has been spent.
We want to estimate the percentage of money from the file that was wrongfully spent, the error rate.

This very tiny package has the function, drawsneeded().
The function gives an estimate of the number of monetary unit
draws needed
to establish with some certainty that the error rate is below a certain threshold.

This is only a good estimation.
Due to randomness, the actual number of monetary unit draws needed
might be either to
low or to high.

The package also has the function draws_needed_plot() which calls drawsneeded()
and shows the result in a plot.

## Installation
You can install the development version of drawsneeded from [GitHub](https://github.com/) with:
```{r install, message=TRUE, warning=TRUE, echo=TRUE}
if (file.exists("/home/crist-jan/R/x86_64-pc-linux-gnu-library/4.5/drawsneeded")) {
  # We are executing on the author machine, use the development version available there.
  loadNamespace("drawsneeded")
} else {
  # Use the github version.
  if (!requireNamespace("pak", quietly = TRUE)) {
    install.packages("pak")
  }
  pak::pak("cfjdoedens/drawsneeded")
}
```

## Example: 0.1 percent error expected

Suppose you know from previous experience
that a small error rate might exist in the mass of monetary statements.
You estimate the error rate to be no more than 0.1 percent.
So you set expected_error_rate to 0.001.
You need a result with 95% certainty.
So you set certainty to 0.95, You decide that the maximum number of samples you are willing to
draw and check is 500.
So you set max_n to 500.

```{r example-tiny-percent-errors-expected}
library(drawsneeded)
drawsneeded(expected_error_rate = 0.001, cert = 0.95, allowed_error_rate = 0.01, max_n = 500)
```
The conclusion is that you will need at least 365 samples.

In a picture this looks like:
```{r plot_example_0.1_percent_error_expected}
library(drawsneeded)
drawsneeded_plot(expected_error_rate = 0.001, cert = 0.95, allowed_error_rate = 0.01, max_n = 500)
```

## Example: no error expected

However, you might expect to see no error at all. Then you could
calculate as follows:
```{r example-no-errors-expected}
library(drawsneeded)
drawsneeded(expected_error_rate = 0.0, cert = 0.95, allowed_error_rate = 0.01, max_n = 500)
```

So probably 298 samples is sufficient.

## Still TODO

* Make drawsneeded() work nicely with vector parameters that have length > 1.
* Handle case where only integer values of k are possible. For example when k > 1 -
cert, then round k up. Otherwise round k down.
* Add extra margin, so extra draws, by taking into account the standard deviation.
* drawsneeded_plot():If possible, depict situation when too many draws are needed.
* drawsneeded_plot() depict as multiple chance graphs different postulated
values for k, in one picture.
