---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!--README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/cfjdoedens/drawsneeded/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cfjdoedens/drawsneeded/actions/workflows/R-CMD-check.yaml)

[![Codecov test
coverage](https://codecov.io/gh/cfjdoedens/drawsneeded/graph/badge.svg)](https://app.codecov.io/gh/cfjdoedens/drawsneeded)

<!-- badges: end -->

```{r init, message=TRUE, warning=TRUE, include=FALSE, paged.print=TRUE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.path = "man/figures/README-",
out.width = "100%"
)
```

# drawsneeded

## Introduction

Supports monetary unit sampling. In monetary unit sampling we have a
file of monetary statements. In general this concerns money that has
been spent. We want to estimate the percentage of money from the file
that was wrongfully spent, the error rate.

This package has the function 

* drawsneeded(expected_error_rate, allowed_error_rate, cert, ...)

The function
gives an estimate of the number of monetary unit draws needed to
establish with some certainty, _cert_, that the error rate is below a certain
threshold, _allowed_error_rate_. The assumption of drawsneeded() is
that each drawn statement contains 100 x _expected_error_rate_ percent
error. Most often, in real life, most monetary statements will contain no errors at all, and some statements
will have large errors, or are even totally wrong.
Still, we suppose that this assumption leaves the estimates
of drawsneeded() realistic.

Due to randomness, the actual number of
monetary unit draws needed might be smaller or bigger. 
Also, it might turn out, that the file has a higher error rate than 
allowed_error_rate. In that case no amount of draws will suffice
to prove that the error rate is below allowed_error_rate.

The package has a couple of plotting functions:

* drawsneeded_plot()
* margin_plot_varying_expected_error_rate()
* margin_plot_varying_allowed_error_rate()
* margin_plot_varying_cert()
* combined_plot()

These should give the user insight in the relation between number
of draws taken and resulting estimates about the error rate in the
file that is audited. 

## Installation

You can install the development version of drawsneeded from
[GitHub](https://github.com/) with:

```{r install, message=TRUE, warning=TRUE, echo=TRUE}
if (file.exists("/home/crist-jan/R/x86_64-pc-linux-gnu-library/4.5/drawsneeded")) {
  # We are executing on the author machine, use the development version available there.
  library("drawsneeded")
  print("using files directly from author of package; not from github")
} else {
  # Use the github version.
  if (!requireNamespace("pak", quietly = TRUE)) {
    install.packages("pak")
  }
  pak::pak("cfjdoedens/drawsneeded")
  print("using files from github")
}
```
## Example: 0.1 percent error expected

### expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = 0.95, max_n = 500

Suppose you know from previous experience that a small error rate might
exist in the mass of monetary statements. You estimate the error rate to
be no more than 0.1 percent of the total mass in money. So you set
expected_error_rate to 0.001. You need a result with 95% certainty. So
you set certainty to 0.95, You decide that the maximum number of samples
you are willing to draw and check is 500. So you set max_n to 500.

```{r tiny-errors-num}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

The conclusion is that you will need at least 365 samples.

In a picture this looks like:

```{r tiny-errors-plot}
drawsneeded_plot(expected_error_rate = 0.001,  allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

We can further analyse the situation by varying over expected_error_rate, or over allowed_error_rate or over cert.
See the following three plots.

### Varying over the expected error rate

First we vary over the expected error rate:

```{r tiny-errors-vary-eer-plot}
margin_plot_varying_expected_error_rate(allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

We see that as expected_error_rate goes near to allowed_error_rate, the
number of draws needed rises.

We can also study this in a non graphical way:

```{r tiny-errors-vary-eer-num}
drawsneeded(expected_error_rate = seq(from = 0.0, by = 0.001, to = 0.009), allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

Again we see that for an expected error rate of 0.001, we need 365 draws.

### Varying over the allowed error rate

```{r tiny-errors-vary-aer-plot}
margin_plot_varying_allowed_error_rate(expected_error_rate = 0.001, cert = 0.95, max_n = 500)
```

We see that as the
allowed_error_rate moves away from the expected_error_rate, the number of
draws falls sharply.

Non graphically this looks like (note the 365 computed draws needed for allowed error rate 0.01):

```{r tiny-errors-vary-aer-num1}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = seq(from = 0.002, by = 0.001, to = 0.01), cert = 0.95, max_n = 500)
```

and,

```{r tiny-errors-vary-aer-num2}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = seq(from = 0.02, by = 0.01, to = 0.1), cert = 0.95, max_n = 500)
```
and,
```{r tiny-errors-vary-aer-num3}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = seq(from = 0.2, by = 0.1, to = 0.9), cert = 0.95, max_n = 500)
```

### Varying over the certainty

Finally we can also vary over the certainty we apply.

```{r tiny-errors-vary-cert-plot}
margin_plot_varying_cert(expected_error_rate = 0.001, allowed_error_rate = 0.01, max_n = 500)
```

As might be expected: if we can do with less certainty, then we need less draws.

In raw numbers this looks like:

```{r tiny-errors-vary-cert-num}
drawsneeded(expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = seq(from = 0.50, by = 0.05, to = 0.95), max_n = 500)
```

Note that for cert 0.95, we get again 365 draws needed.

### All plots bundled in one picture

We can call combined_plots() to get all four plots in one picture. However, it is does not look
very slick here due to lack of canvas space.

```{r tiny-errors-all-plots}
combined_plots(expected_error_rate = 0.001, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```



## Example: no error expected

### expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500

You might expect to see no error at all in the to be audited file of monetary statements. Then you could
calculate as follows:

```{r zero-errors-num}
drawsneeded(expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

So then 298 samples is sufficient.

In a picture this looks like:

```{r zero-errors-plot}
drawsneeded_plot(expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```

Again, we can further analyse the situation by varying over expected_error_rate, or over allowed_error_rate or over cert. As we have above already studied varying the expected_error_rate,
we will not repeat that here.

### Varying over the allowed error rate

```{r zero-errors-vary-aer-plot}
margin_plot_varying_allowed_error_rate(expected_error_rate = 0, cert = 0.95, max_n = 500)
```

We see that as the
allowed_error_rate moves away from the expected_error_rate, the number of
draws falls sharply.

Non graphically this looks like:

```{r zero-errors-vary-aer-num1}
drawsneeded(expected_error_rate = 0, allowed_error_rate = seq(from = 0.001, by = 0.001, to = 0.01), cert = 0.95, max_n = 500)
```

and,

```{r zero-errors-vary-aer-num2}
drawsneeded(expected_error_rate = 0, allowed_error_rate = seq(from = 0.02, by = 0.01, to = 0.1), cert = 0.95, max_n = 500)
```
and,
```{r zero-errors-vary-aer-num3}
drawsneeded(expected_error_rate = 0, allowed_error_rate = seq(from = 0.2, by = 0.1, to = 0.9), cert = 0.95, max_n = 500)
```

### Varying over the certainty

Finally we can also vary over the certainty we apply.

```{r zero-errors-vary-cert-plot}
margin_plot_varying_cert(expected_error_rate = 0, allowed_error_rate = 0.01, max_n = 500)
```

As might be expected: if we can do with less certainty, then we need less draws.

In raw numbers this looks like:

```{r zero-errors-vary-cert-num}
drawsneeded(expected_error_rate = 0, allowed_error_rate = 0.01, cert = seq(from = 0.50, by = 0.05, to = 0.95), max_n = 500)
```

We recognize the number 298 as the number of draws needed for cert 0.95.

### All plots bundled in one picture

We call combined_plots() to get all four plots in one picture. Again, it is does not look
very nice here, because of the cramped space. 

```{r zero-errors-all-plots}
combined_plots(expected_error_rate = 0, allowed_error_rate = 0.01, cert = 0.95, max_n = 500)
```



## Still TODO

-   Handle case where only integer values of k are possible. For example
    when k \> 1 - cert, then round k up. Otherwise round k down.
-   Possibly, add extra margin, so extra draws, by taking into account
    the standard deviation.
-   There is something strange: you pretend to know something about the
    file to be audited, i.e. expected_error_rate \> 0, but you do not
    want to use that info in the estimation, or is that correct as this
    only concerns the planning phase? I get a bit confused.
